---
title: "CaseStudy2_domicia.rmd"
author: "Domicia Herring"
date: "03/27/2021"
output: html_document
editor_options: 
  chunk_output_type: console
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(
	echo = TRUE,
	message = FALSE,
	warning = FALSE,
	comment = NULL
)

```

## 1. Study Background Summary 

  DDSAnalytics is an analytics company that specializes in talent management solutions for Fortune 100 companies. DDSAnalytics has taken on a contract with Frito Lay to tackle predicting employee turnover. Analyzing an existing employee dataset of 870 unique observations with 36 categories, several visualization and models were generated to determine the top three factors leading to attrition.

```{r packages, include=FALSE}
# Load APIs
library(stringr)
library(tidyr)
library(tidyverse)
library(class)
library(caret)
library(e1071)
library(plotly)
library(randomForest)
library(readxl)

## Data Analysis


```


##  2. Importing input data

```{r data}


library(readr)
employeeData <- read.csv("CaseStudy2-data.csv")



# list rows of data that have missing values
employeeData[!complete.cases(employeeData),]

```

  - There are no NA values within the dataset.


```{r}
#Check the data type of the variables in the file
glimpse(employeeData)
```


  - All data types are correct for each variable.


## 3.Exploratory Data Analysis

### 1. Graphs by Job Satistifaction


     

     
### Display Employees Per Job Satistifaction by Gender

```{r}
# count and percent
Gender_Job_table = employeeData %>% 
  count(Gender, JobSatisfaction) %>% 
  group_by(Gender) %>% 
  mutate(proportion = n/sum(n))
Gender_Job_table

#Create visulization
JS_Gender = Gender_Job_table %>% 
  ggplot(aes(x = JobSatisfaction, y = proportion, fill = Gender)) +
  geom_col(show.legend = TRUE, position = "dodge")+
  ggtitle("Employees  Job Satistifaction by Gender") + xlab("Job Satisfaction") + ylab("Proportion")
  
# Display Employees Per Job Satistifaction by Gender
ggplotly(JS_Gender)

```


   
     
### Display Employees Per Job Satisfaction by Age

```{r}
 #count and percent
Age_Job_table = employeeData %>% 
  count(Age, JobSatisfaction) %>% 
  group_by(Age) %>% 
  mutate(proportion = n/sum(n))
Age_Job_table

#Create visualization
JS_Age = Age_Job_table %>% 
  ggplot(aes(x = JobSatisfaction, y = proportion, fill = Age)) +
  geom_col(show.legend = TRUE, position = "dodge")+
  ggtitle("Employees  Job Satistifaction by Age") + xlab("Job Satisfaction") + ylab("Proportion")

#Display Employees Per Job Satisfaction by Age
ggplotly(JS_Age)
  

```

     
### Display Employees Per Overtime by Attrition


```{r}
#count and percent
AttrOT_table = employeeData %>% 
  count(Attrition, OverTime) %>% 
  group_by(Attrition) %>% 
  mutate(proportion = n/sum(n))
AttrOT_table

#Create visualization
Attr_OT = AttrOT_table %>% 
  ggplot(aes(x = OverTime, y = proportion, fill = Attrition)) +
  geom_col(show.legend = TRUE, position = "dodge")+
  ggtitle("Employees  Overtime by Attrition") + xlab("Overtime") + ylab("Proportion")

#Display Employees Per Overtime by Attrition
ggplotly(Attr_OT)


```

### Display Employees Attrition By Job Role

```{r}
#count and percent
AttrJob_table = employeeData %>% 
  count(JobRole,Attrition) %>% 
  group_by(JobRole) %>% 
  mutate(proportion = n/sum(n))
AttrJob_table

# Create visualization 
Attr_Job = AttrJob_table %>% 
  ggplot(aes(x = Attrition, y = proportion, fill = JobRole)) +
  geom_col(show.legend = TRUE, position = "dodge")+
  ggtitle("Employees Attrition by Job Role") + xlab("Job Role") + ylab("Proportion")

#Display Employees Attrition By Job Role
ggplotly(Attr_Job)


```


### Display Employees Overtime By Job Role


```{r}
#count and percent
OTJob_table = employeeData %>% 
  count(JobRole, OverTime) %>% 
  group_by(JobRole) %>% 
  mutate(proportion = n/sum(n))
OTJob_table

#Create Visualization
OTJob = OTJob_table %>% 
  ggplot(aes(x = OverTime, y = proportion, fill = JobRole)) +
  geom_col(show.legend = TRUE, position = "dodge")+
  ggtitle("Employees  Overtime by Job Role") + xlab("Overtime") + ylab("Proportion")

#Display Employees Overtime By Job Role
ggplotly(OTJob)

```

### Display Employees Attrition By Age

```{r}
#count and percent
AttrAge_table = employeeData %>% 
  count(Age, Attrition) %>% 
  group_by(Age) %>% 
  mutate(proportion = n/sum(n))
AttrAge_table

#Create visualization 
AttrAge = AttrAge_table %>% 
  ggplot(aes(x = Attrition, y = proportion, fill = Age)) +
  geom_col(show.legend = TRUE, position = "dodge")+
  ggtitle("Employees  Attrition By Age") + xlab("Attrition") + ylab("Proportion")

#Display Employees Attrition By Age
ggplotly(AttrAge)

```
### Display Employees Per Attrtiton by Monthly Income

```{r}

# count and percent
Attr_MI_table = employeeData %>% 
  count(MonthlyIncome, Attrition) %>% 
  group_by(MonthlyIncome) %>% 
  mutate(proportion = n/sum(n))
Attr_MI_table

# visualization
Attr_MI = Attr_MI_table %>% 
  ggplot(aes(x = Attrition, y = proportion, fill = MonthlyIncome)) +
  geom_col(show.legend = TRUE, position = "dodge")+
  ggtitle("Employees  Attrition by Monthly Income") + xlab("Attrition") + ylab("Proportion")
  
#Display Employees Per Attrition by Monthly Income
ggplotly(Attr_MI)
  


```

## 4. Modeling

### Model 1: Naive Bayes



```{r}

# data prep
df = employeeData
summary(df)

df = df %>% 
  dplyr::select(-ID, - EmployeeCount, - EmployeeNumber, -Over18)  %>% 
  mutate_if(is.character, factor)


## Create train and test sets
set.seed(120)
trainIndices = sample(seq(1, nrow(df), by = 1),(.7*nrow(df)))
trainData = df[trainIndices,]
testData = df[-trainIndices,]

# Naive Bayes model
model.nb = naiveBayes(Attrition ~ ., data = df, positive = "Yes")

# model summary
summary(model.nb )
model.nb$apriori


# Prediction
predData  = predict(model.nb, testData)
table(predData)
predData  = factor(predData)

# Accuracy
confusionMatrix(predData,testData$Attrition, positive = "Yes")

```

  The model requirements were atleast a 60% sensitivity and specificity for the training and validation set. 
  The model met these requirements with a 86.97% accuracy, 67.57% sensitivity and 90.18%.
  
  
```{r random forest}

# data prep
df = employeeData
summary(df)

df = df %>% 
  dplyr::select(-ID, - EmployeeCount, - EmployeeNumber, -Over18)  %>% 
  mutate_if(is.character, factor)

# Random Forest method


model.rf = randomForest(Attrition ~ ., ntree = 100, keep.forest=FALSE,
                  data = df,
                   importance = TRUE)
```


```{r model summary}
model.rf

# Importance Variable Plot

Imp_Var = varImp(model.rf)
(varImpPlot(model.rf))


```
  

```{r}
## Load the validation data set that does not include Attrition


valData = read.csv("CaseStudy2CompSet No Attrition.csv")
#View(valData)

## Export the predicted results from my model into a .csv file for submission
validationPrediction <- predict(model.nb, valData) 
table(validationPrediction)

# create output dataset
output = valData %>% select(ID)
output = output %>% 
  mutate(Attrition = validationPrediction)


# save dataset as a csv
write.csv(output, file = "Case2PredictionsDHerring Attrition.csv", row.names = FALSE)

```
Based on the model to the right, it may be predicted that the top three contributing factors to attrition at Frito Lay are Monthly Income, Overtime and Age.

### Model 2

```{r}
## Run a linear regression with the cleaned data set with Monthly Income for Validation Requirement
fitMonthlyIncome = lm(MonthlyIncome ~ ., data = df)
summary(fitMonthlyIncome)


## Load the validation data set that does not include MonthlyIncome


valData.MI = read_excel("CaseStudy2CompSet No Salary.xlsx")
## Export the predicted results from my model into a .csv file for submission
validationPrediction.MI = predict(fitMonthlyIncome, valData.MI) 


output2 = valData.MI %>% select(ID)
output2 = output2 %>% 
  mutate(MonthlyIncome = validationPrediction.MI)

write.csv(output2, file = "Case2PredictionsDHerring Salary.csv", row.names = FALSE)
```

The model met the required RMSE of less than $3,000 with a RMSE of $1,057.




     